SSwAB-IT Phase 1 Analysis

### 2024-06-28 Jordana Hoppe/SSwAB-IT/phase_1_analysis.qmd
additional subgroup analyses

```{r data cleaning}
#| include: false
# Import from REDCap
unlockREDCap(c(rcon = "SSwAB-IT (New)"),
  keyring = "API_KEYs",
  envir = 1,
  url = "https://redcap.ucdenver.edu/api/"
)
exportBulkRecords(list(db = rcon), envir = 1)
# Exclude
exclude <- c("01-030_W", "01-038_W", "02-0038", "02-0039", "04-001")
# Combine collection results
df <- full_join(db_cos_collection_results, db_sos_collection_results)
# Add dates to demographics
db_demographics <- left_join(db_demographics,
  df %>% select(record_id, cos_result_date),
  by = join_by(record_id)
)
db_demographics$age_at_cos <- as.numeric(difftime(db_demographics$cos_result_date,
  db_demographics$dob,
  units = "days"
)) / 365.25
db_demographics$age_group <- cut(db_demographics$age_at_cos, c(-Inf, 18, Inf),
  right = F
)
# Format
df <- df %>%
  select(
    record_id,
    staphylococcus_aureus_cos:other_result_cos,
    staphylococcus_aureus_sos:other_result_sos
  ) %>%
  filter(!record_id %in% exclude)
# Remove empty demographics rows
db_demographics <- db_demographics %>%
  filter(
    redcap_event_name == "Clinic Visit (Arm 1: Eligible)",
    !record_id %in% exclude
  )
# Excluded participants
db_medications <- db_medications %>% filter(!record_id %in% exclude)
# Combine race and ethnicity columns
races <- c(
  "American Indian or Alaskan Native", "Asian", "Hawaiian or Pacific Islander",
  "Black or African American", "White", "Unknown", "Other"
)
ethnicities <- c("Hispanic", "Non-Hispanic", "Unknown/Not Reported")
db_demographics$Race <- apply(db_demographics, 1, function(r) {
  w <- which(r[grep("race", names(r))] == "Checked")
  if (length(w) == 1) {
    return(races[w])
  } else if (length(w) > 1) {
    return("Multiple Races")
  } else if (length(w) == 0) {
    return(NA)
  }
})
db_demographics$Ethnicity <- apply(db_demographics, 1, function(r) {
  w <- which(r[grep("ethnicity", names(r))] == "Checked")
  if (length(w) == 1) {
    return(ethnicities[w])
  } else if (length(w) > 1) {
    return(paste(ethnicities[w], sep = "/"))
  } else if (length(w) == 0) {
    return(NA)
  }
})
# Add groups to other dataframes
db_medications <- left_join(db_medications,
  db_demographics %>% select(record_id, group),
  by = join_by(record_id)
)
# Bug list
bugs <- c(
  "staphylococcus_aureus", "haemophilus_influenza", "pseudomonas_aeruginosa",
  "pseudomonas_non_aero", "s_maltophilia", "burkholderia", "aspergillus", "ntm",
  "other"
)
bug_names <- c(
  "S. aureus", "H. influenza", "P. aeruginosa", "Pseudomonas non-aeruginosa",
  "S. maltophilia", "Burkholderia", "Aspergillus", "NTM", "Other"
)
# Convert yes/no to positive/negative
df[, paste0(bugs, "_cos")] <- lapply(df[, paste0(bugs, "_cos")], factor,
  levels = c("No", "Yes"), labels = c("Negative", "Positive")
)
df[, paste0(bugs, "_sos")] <- lapply(df[, paste0(bugs, "_sos")], factor,
  levels = c("No", "Yes"), labels = c("Negative", "Positive")
)
# Create overall columns (any bug detected in sample)
df$overall_cos <- apply(df, 1, function(r) {
  any(r[paste0(bugs, "_cos")] == "Positive")
})
df$overall_cos <- factor(df$overall_cos,
  levels = c(F, T), labels = c("Negative", "Positive")
)
df$overall_sos <- apply(df, 1, function(r) {
  any(r[paste0(bugs, "_sos")] == "Positive")
})
df$overall_sos <- factor(df$overall_sos,
  levels = c(F, T), labels = c("Negative", "Positive")
)
# Add overall to list
bugs <- c("overall", bugs)
bug_names <- c("Across all bugs", bug_names)
bug_list <- as.list(bug_names)
names(bug_list) <- bugs
# Add group to dataframe
df <- left_join(df, db_demographics %>% select(record_id, group),
  by = join_by(record_id)
)
# Feasibility
db_feasibility <- db_feasibility %>% filter(!record_id %in% exclude)
```

@@

>>>>

```{r data cleaning}
#| include: false
# Import from REDCap
unlockREDCap(c(rcon = "SSwAB-IT (New)"),
  keyring = "API_KEYs",
  envir = 1,
  url = "https://redcap.ucdenver.edu/api/"
)
exportBulkRecords(list(db = rcon), envir = 1)
# Exclude
exclude <- c("01-030_W", "01-038_W", "02-0038", "02-0039", "04-001")
# Combine collection results
df <- full_join(db_cos_collection_results, db_sos_collection_results)
# Add dates to demographics
db_demographics <- left_join(db_demographics,
  df %>% select(record_id, cos_result_date),
  by = join_by(record_id)
)
db_demographics$age_at_cos <- as.numeric(difftime(db_demographics$cos_result_date,
  db_demographics$dob,
  units = "days"
)) / 365.25
db_demographics$age_group <- cut(db_demographics$age_at_cos,
  c(-Inf, 10, 18, 30, Inf),
  right = F
)
# Format
df <- df %>%
  select(
    record_id,
    staphylococcus_aureus_cos:other_result_cos,
    staphylococcus_aureus_sos:other_result_sos
  ) %>%
  filter(!record_id %in% exclude)
# Remove empty demographics rows
db_demographics <- db_demographics %>%
  filter(
    redcap_event_name == "Clinic Visit (Arm 1: Eligible)",
    !record_id %in% exclude
  )
# Excluded participants
db_medications <- db_medications %>% filter(!record_id %in% exclude)
# Combine race and ethnicity columns
races <- c(
  "American Indian or Alaskan Native", "Asian", "Hawaiian or Pacific Islander",
  "Black or African American", "White", "Unknown", "Other"
)
ethnicities <- c("Hispanic", "Non-Hispanic", "Unknown/Not Reported")
db_demographics$Race <- apply(db_demographics, 1, function(r) {
  w <- which(r[grep("race", names(r))] == "Checked")
  if (length(w) == 1) {
    return(races[w])
  } else if (length(w) > 1) {
    return("Multiple Races")
  } else if (length(w) == 0) {
    return(NA)
  }
})
db_demographics$Ethnicity <- apply(db_demographics, 1, function(r) {
  w <- which(r[grep("ethnicity", names(r))] == "Checked")
  if (length(w) == 1) {
    return(ethnicities[w])
  } else if (length(w) > 1) {
    return(paste(ethnicities[w], sep = "/"))
  } else if (length(w) == 0) {
    return(NA)
  }
})
# Add groups to other dataframes
db_medications <- left_join(db_medications,
  db_demographics %>% select(record_id, group),
  by = join_by(record_id)
)
# Bug list
bugs <- c(
  "staphylococcus_aureus", "haemophilus_influenza", "pseudomonas_aeruginosa",
  "pseudomonas_non_aero", "s_maltophilia", "burkholderia", "aspergillus", "ntm",
  "other"
)
bug_names <- c(
  "S. aureus", "H. influenza", "P. aeruginosa", "Pseudomonas non-aeruginosa",
  "S. maltophilia", "Burkholderia", "Aspergillus", "NTM", "Other"
)
# Convert yes/no to positive/negative
df[, paste0(bugs, "_cos")] <- lapply(df[, paste0(bugs, "_cos")], factor,
  levels = c("No", "Yes"), labels = c("Negative", "Positive")
)
df[, paste0(bugs, "_sos")] <- lapply(df[, paste0(bugs, "_sos")], factor,
  levels = c("No", "Yes"), labels = c("Negative", "Positive")
)
# Create overall columns (any bug detected in sample)
df$overall_cos <- apply(df, 1, function(r) {
  any(r[paste0(bugs, "_cos")] == "Positive")
})
df$overall_cos <- factor(df$overall_cos,
  levels = c(F, T), labels = c("Negative", "Positive")
)
df$overall_sos <- apply(df, 1, function(r) {
  any(r[paste0(bugs, "_sos")] == "Positive")
})
df$overall_sos <- factor(df$overall_sos,
  levels = c(F, T), labels = c("Negative", "Positive")
)
# Add overall to list
bugs <- c("overall", bugs)
bug_names <- c("Across all bugs", bug_names)
bug_list <- as.list(bug_names)
names(bug_list) <- bugs
# Add groups to dataframe
df <- left_join(df, db_demographics %>% select(record_id, group, age_group),
  by = join_by(record_id)
)
df <- left_join(df, db_medications, by = join_by(record_id, group))
# Feasibility
db_feasibility <- db_feasibility %>% filter(!record_id %in% exclude)
```

@@

## Age group

```{r results='asis'}
invisible(lapply(names(bug_list), function(b) {
  cat("\n")
  cat("###", as.character(bug_list[b]))
  cat("\n")
  cat("####", "Agreement by age group")
  cat("\n")
  agreement_test(b, group = "age_group")
  cat("\n")
}))
```

## Sex

```{r results='asis'}
invisible(lapply(names(bug_list), function(b) {
  cat("\n")
  cat("###", as.character(bug_list[b]))
  cat("\n")
  cat("####", "Agreement by gender assigned at birth")
  cat("\n")
  agreement_test(b, group = "age_group")
  cat("\n")
}))
```

## Azithromycin

```{r results='asis'}
invisible(lapply(names(bug_list), function(b) {
  cat("\n")
  cat("###", as.character(bug_list[b]))
  cat("\n")
  cat("####", "Agreement by Azithromycin status")
  cat("\n")
  agreement_test(b, group = "azithromycin")
  cat("\n")
}))
```

## Inhaled Tobramycin

```{r results='asis'}
invisible(lapply(names(bug_list), function(b) {
  cat("\n")
  cat("###", as.character(bug_list[b]))
  cat("\n")
  cat("####", "Agreement by inhaled Tobramycin status")
  cat("\n")
  agreement_test(b, group = "inhaled_tobramycin")
  cat("\n")
}))
```

## Inhaled Aztreonam

```{r results='asis'}
invisible(lapply(names(bug_list), function(b) {
  cat("\n")
  cat("###", as.character(bug_list[b]))
  cat("\n")
  cat("####", "Agreement by inhaled Aztreonam status")
  cat("\n")
  agreement_test(b, group = "inhaled_aztreonam")
  cat("\n")
}))
```

## Inhaled Colisitin

```{r results='asis'}
invisible(lapply(names(bug_list), function(b) {
  cat("\n")
  cat("###", as.character(bug_list[b]))
  cat("\n")
  cat("####", "Agreement by inhaled Colisitin status")
  cat("\n")
  agreement_test(b, group = "inhaled_colisitin")
  cat("\n")
}))
```

## Inhaled Vancomycin

```{r results='asis'}
invisible(lapply(names(bug_list), function(b) {
  cat("\n")
  cat("###", as.character(bug_list[b]))
  cat("\n")
  cat("####", "Agreement by inhaled Vancomycin status")
  cat("\n")
  agreement_test(b, group = "inhaled_vancomycin")
  cat("\n")
}))
```

## Other inhaled antibiotics

```{r results='asis'}
invisible(lapply(names(bug_list), function(b) {
  cat("\n")
  cat("###", as.character(bug_list[b]))
  cat("\n")
  cat("####", "Agreement by other inhaled antibiotics status")
  cat("\n")
  agreement_test(b, group = "other_inhaled_antibiotics")
  cat("\n")
}))
```

## Chronic oral antibiotic

```{r results='asis'}
invisible(lapply(names(bug_list), function(b) {
  cat("\n")
  cat("###", as.character(bug_list[b]))
  cat("\n")
  cat("####", "Agreement by chronic oral antibiotic status")
  cat("\n")
  agreement_test(b, group = "chronic_oral_antibiotic")
  cat("\n")
}))
```

## CFTR modulator status

```{r results='asis'}
invisible(lapply(names(bug_list), function(b) {
  cat("\n")
  cat("###", as.character(bug_list[b]))
  cat("\n")
  cat("####", "Agreement by CFTR modulator status")
  cat("\n")
  agreement_test(b, group = "cftr_modulator")
  cat("\n")
}))
```
